---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rcits

<!-- badges: start -->
<!-- badges: end -->

The goal of this package, given a causal and invertible time series model, is two-fold: (i) given innovations, simulate observations via `inn2ts`, (ii) given observations, return innovations via `ts2inn`.
The first is natural in simulating time series.
The second is useful for getting fitted residuals where the parameter could have been estimated on a different dataset compared to the dataset on which residuals are computed on.

This package implements the two methods for ARIMA and GARCH models.

## Installation

You can install the development version of simults like so:

``` r
pak::pkg_install("leonfernandes/rcits")
```

## Simulate Data

We demonstrate the steps to simulate an AR(1) below.

```{r sim-ar1}
library(rcits)

mdl <- make_arima(phi=0.3, theta=0, delta=0)
z <- rnorm(200)
x <- inn2ts(mdl, inn=z)
plot(x, type='l', main="Simulated AR(1)")
```

## Exact Residuals

If we apply `obs2inn` using the true model parameters, we should recover the original innovations.
This is verified below.

```{r fitted_resid}
z0 <- ts2inn(mdl, x) # Exact residuals
print(tail(z))
print(tail(z0$inn))
```

## Fitted Residuals

Fit an AR(1) model on first half of the data and compute residuals on all the data.

```{r fitted-plot}
phi_hat <- stats::ar(x$ts[1:50], order.max=1)$ar
fit_mdl <- make_arima(phi=phi_hat, theta=0, delta=0)
z_hat <- ts2inn(fit_mdl, x)
print(tail(z_hat$inn))
plot(z_hat, type='l', main="Fitted residuals from AR(1)")
```