---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# simults

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/leonfernandes/simults/branch/master/graph/badge.svg)](https://app.codecov.io/gh/leonfernandes/simults?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/simults)](https://CRAN.R-project.org/package=simults)
[![BioC status](http://www.bioconductor.org/shields/build/release/bioc/simults.svg)](https://bioconductor.org/checkResults/release/bioc-LATEST/simults)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of simults is to simulate time series from various time series models.

## Installation

You can install the development version of simults like so:

``` r
pak::pkg_install("leonfernandes/simults")
```

## Simulate Data

We demonstrate how to simulate an AR(1) below.

```{r sim-ar1}
library(simults)
# simulate from an AR(1) model
mdl <- make_arima(phi=0.3, theta=0, delta=0)
x <- simults(mdl, nsim=100, innov=rnorm(200))
plot(x, type='l', main="Simulated AR(1)")
```
Currently this package supports ARIMA and GARCH time series models from the *smpspltools* package.
Burn-in is done implicitly where sufficient innovations are assumed to be supplied.

## Fitted Residuals

Fit an AR(1) model on first half of the data and compute residuals on all the data.
```{r fitted-plot}
phi_hat <- stats::ar(x$value[1:50], order.max=1)$ar
fit_mdl <- make_arima(phi=phi_hat, theta=0, delta=0)
z <- fitted_resid(fit_mdl, x)
plot(z, type='l', main="Fitted residuals from AR(1)")
```